Clouds and Sensitivities Across a Hierarchy of GFDL CMIP6 Models

### 2020-08-26 Silvers_cfmip.tex
relatively minor textual edits to Silvers_cfmip.tex

The comparisons with observations are made mostly with the \textit{amip} experiment.   The experiments used 
in this paper are all part of the GFDL contribution to CMIP6.  We use the \textit{historical, abrupt-4xCO2, and 1pctCO2}
experiments from the CMIP6 DECK.  In addition we utilize several experiments from the CFMIP contributions 
to CMIP6.  These experiments include \textit{amip, amip-p4K, amip-m4K, amip-future4K, aqua-control} and \textit{aqua-p4K}.

@@

\textit{This paragraph should serve as a bridge between the COSP observations, amip and historical experiments, TOA fluxes, and feedbacks.
While direct comparisons between GCM output and satellite observations are only possible over the recent decades, experiments that use
historically based data and extend over longer time periods can give insight into the relationship between cloud area, radiative fluxes
and the decadal variability of the climate feedback parameter.  The \textit{amip}, \textit{amipPIforcing}, and \textit{historical} experiments are useful
for these purposes.  Pronounced decadal variability of the climate feedback parameter has been shown to be closely tied to 
variations of the SW CRE and the low-level clouds (Gregory and Andrews, 2016; Zhou et al., 2016; Silvers et al., 2018).}
Plotting the cloud fields as a function of latitude and height emphasizes the similarities and differences 
between the various experiments while also allowing for a comparison with the high-, mid-, and low-level clouds 
that have been observed by CALIPSO.  Overall the clouds in the $amip$ and $historical$ simulations are similar to 
each other, as we would expect (Figure \ref{fig:calipso_amip_historical}) given their similar radiative forcing 
and experimental configuration.  The differences between the clouds of 
these two experiments are due to the different SST and sea-ice patterns.  The SST patterns and 
sea ice concentrations in the \textit{amip} experiment are
prescribed based on observations, while those of the \textit{historical} experiment are the result of the 
interactions between the atmosphere and dynamic ocean and sea-ice models.  
Figure \ref{fig:calipso_amip_historical} thus shows how the clouds change 
as a result of SST biases over the observational period.   There are two distinct biases in this case, 
the bias in SST generated from differences between the SST of the coupled model and observations,
%the SST bias  that is due to the difference between the SST generated 
%from the coupled model and observations 
as well as the bias in the cloud field relative to observations that is present in both 
experiments.  As reference, the latitudinal profiles of cloud fraction from the CALIPSO observations that 
were discussed in the previous section are also plotted in  Figure \ref{fig:calipso_amip_historical}.
The overall latitudinal structure of the cloud fraction profiles match well between the model results and 
CALIPSO but the negative cloud bias from AM4 and CM4 is apparent at most latitudes and all three levels.  
There is a particularly large positive bias poleward of about 60N in the low-level clouds.

@@

The most recent generation of models developed at NOAA's GFDL that have participated in CMIP6 have a large range of climate sensitivities (possibly cite Held et al., Winton et al., Dunne et al.).  
One way of quantifying this sensitivity is with the climate feedback parameter, which, for the subset of GFDL models and CMIP6 experiments studied in this paper, varies between -0.46 and -2.1 $\rm{W m^{-2}K^{-1}}$ (Table 1).  
These models (AM4.0 Aquaplanet, AM4.0, CM4.0) simulate several 
distinct geophysical configurations which account for some of the range in values of the climate feedback parameter.  
However, our hypothesis is that the large range of climate sensitivities that are found within this 
group of models is in large part due to variations in the clouds (both their climatology and their feedbacks).

@@

>>>>

The comparisons with observations are made mostly with the \textit{amip} experiment.   The experiments used 
in this paper are all part of the GFDL contribution to CMIP6 and as a result, the \textit{amip} and \textit{historical}
experiments end in the year 2014.  We use the \textit{historical, abrupt-4xCO2, and 1pctCO2}
experiments from the CMIP6 DECK.  In addition we utilize several experiments from the CFMIP contributions 
to CMIP6.  These experiments include \textit{amip, amip-p4K, amip-m4K, amip-future4K, aqua-control} and \textit{aqua-p4K}.  
In order to utilize as much of the available data as possible much of the analysis of the \textit{amip} experiment uses the 
full span of years (1979-2014) despite the fact that the satellite observations do not extend as far back in time.  However, 
some of the analysis for the \textit{amip} and \textit{historical} experiments, in particular for the calculation of a climate feedback parameter over recent years, a shorter period (1996-2014) is used in order to eliminate complications due to radiative forcing in the early 1990's due to volcanic activity.  The relevant time period is stated in the captions of the figures and tables.

@@

The comparisons of cloud fraction in Figures 1-4 between observations and the \textit{amip} experiment show a 
prevailing negative cloud fraction bias across much of the globe for low-, middle-, and upper-level clouds.  Does 
coupling the AM4.0 atmosphere to a dynamic ocean decrease the cloud bias?   
\textit{This paragraph should serve as a bridge between the COSP observations, amip and historical experiments, TOA fluxes, and feedbacks.
While direct comparisons between GCM output and satellite observations are only possible over the recent decades, experiments that use
historically based data and extend over longer time periods can give insight into the relationship between cloud area, radiative fluxes
and the decadal variability of the climate feedback parameter.  The \textit{amip}, \textit{amipPIforcing}, and \textit{historical} experiments are useful
for these purposes.  Pronounced decadal variability of the climate feedback parameter has been shown to be closely tied to 
variations of the SW CRE and the low-level clouds (Gregory and Andrews, 2016; Zhou et al., 2016; Silvers et al., 2018).}
Plotting the cloud fields as a function of latitude and height emphasizes the similarities and differences 
between the various experiments while also allowing for a comparison with the high-, mid-, and low-level clouds 
that have been observed by CALIPSO.  Overall the clouds in the $amip$ and $historical$ simulations are similar to 
each other, as we would expect (Figure \ref{fig:calipso_amip_historical}) given their similar radiative forcing 
and experimental configuration.  The differences between the clouds of 
these two experiments are due to the different SST and sea-ice patterns.  The SST patterns and 
sea ice concentrations in the \textit{amip} experiment are
prescribed based on observations, while those of the \textit{historical} experiment are the result of the 
interactions between the atmosphere and dynamic ocean and sea-ice models.  
Figure \ref{fig:calipso_amip_historical} thus shows how the clouds change 
as a result of SST biases over the observational period.   There are two distinct biases in this case, 
the bias in SST generated from differences between the SST of the coupled model and observations,
%the SST bias  that is due to the difference between the SST generated 
%from the coupled model and observations 
as well as the bias in the cloud field relative to observations that is present in both 
experiments.  As reference, the latitudinal profiles of cloud fraction from the CALIPSO observations that 
were discussed in the previous section are also plotted in  Figure \ref{fig:calipso_amip_historical}.
The overall latitudinal structure of the cloud fraction profiles match well between the model results and 
CALIPSO but the negative cloud bias from AM4 and CM4 is apparent at most latitudes and all three levels.  
There is a particularly large positive bias poleward of about 60N in the low-level clouds.

@@

The most recent generation of models developed at NOAA's GFDL that have participated in CMIP6 have a large range of climate 
sensitivities (possibly cite Winton et al., Dunne et al.) depending on the experiment of interest and the method used for 
calculation.  One way of quantifying this sensitivity is with the climate feedback parameter, which, for the subset of GFDL models and CMIP6 experiments studied in this paper, varies between -0.46 and -2.1 $\rm{W m^{-2}K^{-1}}$ (Table 1).  
These models (AM4.0 Aquaplanet, AM4.0, CM4.0) simulate several 
distinct geophysical configurations which account for some of the range in values of the climate feedback parameter.  
However, our hypothesis is that the large range of climate sensitivities that are found within this 
group of models is in large part due to variations in the clouds (both their climatology and their feedbacks).

@@

\textit{Put the results of AM4.0/CM4.0 in the context of other modeling studies.  The AR5, chapter 7 describes the primary 
cloud feedbacks that are thought to be relevant for future warming.  Are AM4.0/CM4.0 consistent with the expectations from the AR5?  Can we say something about the large cloud fraction in the arctic?  Tie into Kay et al. 2011 and Vavrus et al. (2009 or 2011)}
